---
title: "pproject"
author: "SaiPriya Gourineni"
date: "2022-12-14"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data Collection

```{r}
data <- read.csv("C:/Users/Saipr/Downloads/diabetes.csv")
head(data)
```

Data Summary

```{r}
summary(data)
```

A value of 0 in these columns indicates a missing value since it is illogical.

The 0 value in the following columns or variables is invalid:Glucose,BloodPressure ,SkinThickness ,Insulin and BMI

Since counting them thereafter would be simpler and zeros need to be replaced with appropriate values, it is preferable to replace zeros with nan.

Replacing 0 with Nans

```{r}
data["BloodPressure"][data["BloodPressure"] == 0] <- NA
data["Glucose"][data["Glucose"] == 0] <- NA
data["SkinThickness"][data["SkinThickness"] == 0] <- NA
data["Insulin"][data["Insulin"] == 0] <- NA
data["BMI"][data["BMI"] == 0] <- NA
```

```{r}
summary(data)
```

```{r}
par(mfrow = c(3, 3))
hist(data$BloodPressure)
hist(data$Glucose)
hist(data$SkinThickness)
hist(data$Insulin)
hist(data$BMI)
```

Replacing Nan values with the appropriate median/mean according to the data histogram

```{r}
data$BloodPressure[is.na(data$BloodPressure)] <- round(mean(data$BloodPressure, na.rm = TRUE))
data$Glucose[is.na(data$Glucose)] <- round(mean(data$Glucose, na.rm = TRUE))
data$SkinThickness[is.na(data$SkinThickness)] <- round(median(data$SkinThickness, na.rm = TRUE))
data$Insulin[is.na(data$Insulin)] <- round(median(data$Insulin, na.rm = TRUE))
data$BMI[is.na(data$BMI)] <- round(mean(data$BMI, na.rm = TRUE))
```

```{r}
summary(data)
```

plotting the histograms after imputation:

```{r}
par(mfrow = c(3, 3))
hist(data$BloodPressure)
hist(data$Glucose)
hist(data$SkinThickness)
hist(data$Insulin)
hist(data$BMI)
```

```{r}
count_0 = 0
for (i in data$Outcome) {
  if (i == 0)
    count_0 = count_0 +1
}
count_0
count_1 = length(data$Outcome) - count_0
count_1
```

```{r}
A <- c(count_0, count_1)
  
# Plot the bar chart 
barplot(A, xlab = "0 and 1 Count", ylab = "Count", main ="Outcome count")
```

```{r}
length(data)
length(data$Pregnancies)
```

Correlation Heatmap for cleaned data

```{r}
cormat <- round(cor(data),2)
cormat
```

```{r}
library(reshape2)
 

 
melted_corr_mat <- melt(cormat)
head(melted_corr_mat)
 
# plotting the correlation heatmap
library(ggplot2)
ggplot(data = melted_corr_mat, aes(x=Var1, y=Var2,
                                   fill=value)) +
geom_tile() +
 scale_x_discrete(guide = guide_axis(n.dodge=3))+ 
geom_text(aes(Var2, Var1, label = value),
          color = "black", size = 4)


```

Feature Scaling

```{r}
library(dplyr)
library(caret)
row_labels = data[,9]
data[,1:8]<- as.data.frame(scale(data[,1:8]))
row_labels
```

setting the seed so as to get the same result everytime and splitting the data into train and test

```{r}

set.seed(123)
size <- floor(0.8 *  nrow(data))
train_ind <- sample(seq_len(nrow(data)), size = size)

data_train <- data[train_ind,1:8]
data_test <- data[-train_ind,1:8]

train_labels <- data[train_ind, 9]
test_labels <- row_labels[-train_ind]
length(data$Pregnancies)
data_test
```

```{r}
library(class)

predictions <- knn(train = data_train,
                   test = data_test,
                   cl = train_labels,
                   k= 9)
length(predictions)

```

Confusion Matrix

```{r}


##create confusion matrix
tab <- table( predictions, test_labels ) 
tab
```

```{r}
prediction <- factor(c(0, 0, 1, 1))
test_label <- factor(c(0, 1, 0, 1))
Y      <- c(11, 34, 91, 18)
df <- data.frame(prediction, test_label, Y)
```

```{r}

library(ggplot2)
ggplot(data =  df, mapping = aes(x = test_label, y = prediction)) +
  geom_tile(aes(fill = Y), colour = "white") +
  geom_text(aes(label = sprintf("%1.0f", Y)), vjust = 1) +
  scale_fill_gradient(low = "blue", high = "red") +
  theme_bw() + theme(legend.position = "none")
```

Accuracy

```{r}

##check the accuracy
accuracy <-    function( matriz ){
        sum( diag( x = matriz ) / sum( rowSums( x = matriz )) ) * 100.0
}



print( paste('accuracy ' , round( accuracy( tab ) , digits = 3 )  ) )
```

```{r}
data_test
```

```{r}
plot_predictions <- data.frame(
  data_test$Pregnancies,
  data_test$Glucose,
  data_test$BloodPressure,
  data_test$SkinThickness,
  data_test$Insulin,
  data_test$BMI,
  data_test$DiabetesPedigreeFunction,
  data_test$Age,
  
  predicted = predictions)

colnames(plot_predictions) <- c("data_test$Pregnancies",
                                "data_test$Glucose",
                                "data_test$BloodPressure",
                                "data_test$SkinThickness",
                                "data_test$Insulin",
                                "data_test$BMI",
                                "data_test$DiabetesPedigreeFunction",
                                "data_test$Age",
                                "predicted")
# Visualize the KNN algorithm results.
library(ggplot2)
library(plyr)
require(gridExtra)



p1 <- ggplot(plot_predictions, aes(data_test$Pregnancies,data_test$Glucose, color = predicted, fill = predicted)) + 
  geom_point(size = 5) + 
  geom_text(aes(label=test_labels),hjust=1, vjust=2) +
  ggtitle("Predicted relationship between Pregnancies and Glucose Intake") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none")


p2 <- ggplot(plot_predictions, aes(data_test$BloodPressure, data_test$SkinThickness, color = predicted, fill = predicted)) + 
  geom_point(size = 5) + 
  geom_text(aes(label=test_labels),hjust=1, vjust=2) +
  ggtitle("Predicted relationship between BloodPressure and Skinthickness") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none")

grid.arrange(p1, p2, ncol=1)
```
