---
title: "64060_Assignment 2"
author: "Saipriya Gourineni"
date: "2022-10-02"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Below are the packages used.

```{r}
library(caret)
library(class)
library(ISLR)	
library(dplyr)
library(ggplot2)
library(fastDummies)
library(FNN)
```
##loaded the UniversalBank.csv file with customer data and transform the categorical data to factors.

```{r}
getwd()
setwd("C:/Users/Saipr/Downloads")
BankInfo <- read.csv("C:/Users/Saipr/Downloads/UniversalBank.csv")
BankInfo$Personal.Loan<-factor(BankInfo$Personal.Loan,levels=c('0','1'),labels=c('No','Yes'))
summary(BankInfo)
```

## Data Selection

## Utilizing the data, we should divide the collection into training (60%) and validation (40%). ( ID and Zip for each education level will be transformed Education into three dummy variables).
```{r}
dummy_BankInfo <- dummy_columns(BankInfo, select_columns = 'Education')
m_BankInfo <- select(dummy_BankInfo,Age,Experience,Income,Family,CCAvg,Education_1,Education_2,Education_3,Mortgage,Personal.Loan,Securities.Account,CD.Account,Online,CreditCard)
m_BankInfo <- m_BankInfo %>% relocate(Personal.Loan,.after=last_col())
set.seed(1)
Train_Index <- sample(row.names(m_BankInfo), .6*dim(m_BankInfo)[1])
Val_Index <- setdiff(row.names(m_BankInfo), Train_Index)
Train_Data <- m_BankInfo[Train_Index,]
Validation_Data <- m_BankInfo[Val_Index,]
summary(Train_Data)

```
##Normalizing the Data.

```{r}
columnsare <-c(1,2,3,4,5,9)
BankInfo.norm.df <- m_BankInfo
train.norm.df <- Train_Data
valid.norm.df <- Validation_Data
norm.values <- preProcess(Train_Data[,columnsare], method=c("center","scale"))

train.norm.df[, columnsare] <-predict(norm.values,Train_Data[,columnsare])
valid.norm.df[, columnsare] <-predict(norm.values,Validation_Data[,columnsare])
summary(train.norm.df)
```
##K-NN Model

```{r}
train.knn.predictors <- train.norm.df[, 1:13]
train.knn.success <-train.norm.df[,14]
valid.knn.predictors <- valid.norm.df[, 1:13]
valid.knn.success <-valid.norm.df[,14]
knn.results <- knn (train=train.knn.predictors, test=valid.knn.predictors, cl=train.knn.success, k=1, prob=TRUE)
confusionMatrix(knn.results,valid.knn.success, positive="Yes")
```

## The model is 95.4% accurate.

## k=1

## Below are the characteristics provided: 
Age = 40,Experience = 10,Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1,Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1, and Credit Card = 1. 


```{r}
customertest = data.frame(Age = as.integer(40), Experience = as.integer(10), Income = as.integer(84), Family = as.integer(2), CCAvg = as.integer(2), Education1 = as.integer(0), Education2 = as.integer(1), Education3 = as.integer(0), Mortgage = as.integer(0), Securities.Account = as.integer(0), CD.Account = as.integer(0), Online = as.integer(1), CreditCard = as.integer(1))

#customertest is being loaded to the dataframe.

customer.norm.df <- customertest
customer.norm.df[, columnsare]<-predict(norm.values,customertest[,columnsare])

#normalize the quantitative values
```

## NOW testing with K-NN from earlier.

```{r}
set.seed(400)
customer.knn <- knn(train=train.knn.predictors, test=customer.norm.df,cl=train.knn.success,k=1, prob=TRUE) #calculate knn for customer.
head(customer.knn) 

```

## Tuning using Validation

```{r}
accuracy.df <- data.frame(k = seq(1,14,1), accuracy = rep(0 , 14))


for(i in 1:14){
  knn.pred <- knn(train.knn.predictors,valid.knn.predictors, cl=train.knn.success,k=i)
accuracy.df[i,2] <- confusionMatrix(knn.pred, valid.knn.success)$overall[1]
  }
accuracy.df
plot(x=accuracy.df$k, y=accuracy.df$accuracy, main="Accuracy vs K", xlab="k",ylab="accuracy")
which.max(accuracy.df$accuracy)

```

## Now we will make a table with all of k and their accuracies from 1 to 14.The k balances overfitting and ignoring predictions, and is the accuracy for k is 3.


```{r}
customer.knn3 <- knn(train=train.knn.predictors, test=customer.norm.df,cl=train.knn.success,k=3, prob=TRUE)
head(customer.knn3)

```

## Further test of k = 3

## A confusion matrix of the validation data for k=3 is shown below

```{r}
knn.k3 <- knn(train = train.knn.predictors,test=valid.knn.predictors,cl=train.knn.success,k=3, prob=TRUE)
confusionMatrix(knn.k3,valid.knn.success,)
```

## Repartitioning for a test set

```{r}
set.seed(500)
Train_Index <- sample(row.names(m_BankInfo), .5*dim(m_BankInfo)[1])
Val_Index <- sample(setdiff(row.names(m_BankInfo),Train_Index),.3*dim(m_BankInfo)[1])
Test_Index =setdiff(row.names(m_BankInfo),union(Train_Index,Val_Index))
#load the data
Train_Data <- m_BankInfo[Train_Index,]
Validation_Data <- m_BankInfo[Val_Index,]
Test_Data <- m_BankInfo [Test_Index,]

#normalize the quantitative data
norm.values3 <- preProcess(m_BankInfo[,columnsare], method=c("center", "scale"))
train.norm.df3 = Train_Data
val.norm.df3 = Validation_Data
test.norm.df3 = Test_Data
train.norm.df3[, columnsare] <- predict(norm.values3, Train_Data[, columnsare])
val.norm.df3[, columnsare] <- predict(norm.values3, Validation_Data[, columnsare])
test.norm.df3[, columnsare] <- predict(norm.values3, Test_Data[, columnsare])

#run knn for all 3

knn.train <- knn(train=train.norm.df3[,-14],test=train.norm.df3[,-14],cl=train.norm.df3[,14], k=3, prob=TRUE)
knn.val<- knn(train=train.norm.df3[,-14],test=val.norm.df3[,-14],cl=train.norm.df3[,14],k=3, prob=TRUE)
knn.test<- knn(train=train.norm.df3[,-14],test=test.norm.df3[,-14],cl=train.norm.df3[,14],k=3, prob=TRUE)

#display the confusion matrices

confusionMatrix(knn.train,train.norm.df3[,14], positive="Yes")
confusionMatrix(knn.val,val.norm.df3[,14], positive="Yes")
confusionMatrix(knn.test,test.norm.df3[,14], positive="Yes")
```

